% This is a small sample LaTeX input file (Version of 10 April 1994)
%
% Use this file as a model for making your own LaTeX input file.
% Everything to the right of a  %  is a remark to you and is ignored by LaTeX.

% The Local Guide tells how to run LaTeX.

% WARNING!  Do not type any of the following 10 characters except as directed:
%                &   $   #   %   _   {   }   ^   ~   \

\documentclass{article}
\setcounter{secnumdepth}{0}

% --------------------------- sample ---------------------------------

% \section{Simple Text}          % This command makes a section title.

% Words are separated by one or more spaces.  Paragraphs are separated by
% one or more blank lines.  The output is not affected by adding extra
% spaces or extra blank lines to the input file.

% Double quotes are typed like this: ``quoted text''.
% Single quotes are typed like this: `single-quoted text'.

% Long dashes are typed as three dash characters---like this.

% Emphasized text is typed like this: \emph{this is emphasized}.
% Bold       text is typed like this: \textbf{this is bold}.

% \subsection{A Warning or Two}  % This command makes a subsection title.

% If you get too much space after a mid-sentence period---abbreviations
% like etc.\ are the common culprits)---then type a backslash followed by
% a space after the period, as in this sentence.

% Remember, don't type the 10 special characters (such as dollar sign and
% backslash) except as directed!  The following seven are printed by
% typing a backslash in front of them:  \$  \&  \#  \%  \_  \{  and  \}.
% The manual tells how to make other symbols.

% --------------------------------------------------------------------

\title{\textbf{Companion to the Manual} \\ Celebrating Life as a Catholic Christian}
\date{}


CELEBRATING LIFE AS A CATHOLIC CHRISTIAN

This booklet is a companion to a larger book, a manual titled, Celebrating Life
as a Catholic Christian (CLCC). In a few pages it tells you what the CLCC is: a
process that forms Catholics in the faith through official Church teachings and
helps them respond to their ``Call to Holiness.'' It will also explain CLCC's
special apostolate, reaching out to and helping an estimated 45 million lapsed,
non-practicing Catholics return to Christ and his Church, a mammoth challenge
for the Church.

\section{Small Ecclesial Faith Communities}

The process of formation takes place in Small Ecclesial Faith Communities
(SEFC): Small, a group of 8 to 12 people who meet in someone's home; Ecclesial,
the group is composed of parish members and is approved by the pastor; Faith,
the group is eager to know the faith better, grow in love of God and neighbor,
is earnest about spreading the Good News, and building up the Church to ``full
stature;'' and Community, an assortment of fellow Catholics, who previously knew
each other only by face now come together in loving ``communion'' and bond
through weekly study, sharing, prayer and mutual support.
Jesus' primary reason for coming to earth was to establish his Church, at once
an earthly and heavenly kingdom. He started by forming his own small faith
community of twelve Apostles. Since his day the apostles and their successors
through the centuries have established and depended on small faith communities
for building the kingdom. Historically small faith communities have a missionary
purpose. They are a catalyst for parishes, from which dioceses are
erected. Today they dot every corner of the earth making the Church undeniably
``catholic'' or universal. They are the starting point for giving growth and
renewal to the Church.
Pope John Paul II during his twenty-seven year pontificate preached and promoted
what he called ``Small Ecclesial Faith Communities.'' He added the word
``ecclesial'' to make sure they do not function independently of the
Church. These he saw as an efficient way for doing New Evangelization, the Good
News that he said must be ``new in ardor, new in expression and new in method.''
(Haiti, 1983) The CLCC process is indeed ``new'' in all three categories.

\section{Catholic Identity}

A currently critical issue in the Church is loss of ``Catholic identity.'' The
CLCC process, Celebrating Life as a Catholic Christian, is a way to recover
it. Loss occurs because so many Catholics know so little about the ``mystery of
Church.'' Since Vatican Council II the Popes, in particular John Paul II and
Paul
2
VI, have been urging the People of God to learn more about the Church through
systematic formation drawn from her magisterial and authoritative
teachings. Better understanding of the Church and deeper spirituality result.
Being Catholic means more than being baptized and attending Sunday Mass in some
nearby Catholic church. It means more than pumping a little religion into
children grades K through 8, and after the Sacrament of Confirmation is
conferred seeing them no longer. Formation in Church teachings and holiness must
be ongoing especially for adults, young adults and youth. In the past forty
years most Catholic unfortunately have been given a disastrously inferior
education in the faith. Religious illiteracy among great number of Catholic is
endemic, Asked a question about the faith, the answer is either slow in coming
or embarrassingly inadequate.
Catholic identity means that a Catholic's life style, grounded in solid
theological and moral principles, sets him apart from the crowd. Good character
qualities that are observable also make him or her different. Identity is not a
religious medal dangling on a chain from one's neck. It's attitude, values,
demeanor and conviction, all are centered in a derived from supernatural faith,
hope and charity, humility and wisdom that emanate outward. Both Catholic
Identity and Catholic ``in name only'' are transparent.
To be ``different'' means becoming engaged in the Spirit-filled promise that
Christ made without any reservations, ``Where two or three are gathered together
in my name, there am I in the midst of them.'' (Mt. 18.20) Members of Small
Ecclesial Faith Communities gradually get to the point that they sense Christ's
Spirit is present to them. As they acknowledge that the Holy Spirit is pouring
out an abundance of graces on them, so their sense of Catholic identity grows
stronger. Gathering together week after week for formation in holiness and truth
marks each individual of the group as ``different,'' a complement not a
criticism.
Grace is the ``living water'' that satisfies the soul's thirst for knowledge and
holiness as Jesus stated to the Samaritan woman at
3
the well. (Cf. John 4.10) The CLCC formation process brings about conversion
from a run of the mill Catholic to one who identifies genuinely as
Catholic. What the CLCC formation process does is raise consciousness among
God's People to the Catholic Identity crisis that exists in the Church and in
our culture. It invites them not to ignore it but be willing to do something
about it: begin Celebrating Life as a Catholic Christian.

\section{St. Peter's Words}

St. Peter in his First Letter (2.9) identifies newly established small Christian
communities in the early church in this way. You are ``a chosen race, a
community of priest-kings, a consecrated nation, a people God has made his own
to proclaim his wonders.'' By these words St. Peter is stating how essential
small Christian faith communities are for Celebrating Life as a Catholic
Christian.
In the same letter St. Peter uses another analogy likening Small Faith
Communities to living stones. ``Like living stones,'' he writes, ``let
yourselves be built into a spiritual house to be a holy priesthood to offer
spiritual sacrifices acceptable to God through Jesus Christ.'' (2.5) St. Peter
envisioned the Church as a ``spiritual house,'' the House of God, built with
``living stones,'' small faith communities, growing into bigger spiritual
houses, and ultimately becoming the Universal Church. In the two texts one can
see Catholic Identity, ``living stones'' starting with Small Faith Communities
and growing into the corporate Body of Christ. The Church's beginning on
Pentecost Sunday, its increase and renewal over the centuries, starts now as
then through Small Faith Communities. Both of St. Peter's statements hold a
wealth of meaning. To comprehend the depth of their meaning is food for a
lifetime of reflection.
4

\section{First Considerations of the CLCC Process}

Jesus' parting words to his Apostles before ascending into heaven were: ``Go,
therefore, make disciples of all nations... teaching the to observe all that I
have commanded you. And Behold, I am with you always, until the end of the
age.'' (Mt. 2.19- 20) This command is Christ's blueprint for building up his
Church. The Vatican Council's Decree on the Apostolate of the Laity states that
the laity have a ``right and duty to exercise the apostolate... (having) their
own proper roles in building up the Church.'' (25) Practically this means the
laity can and should share the function of building up the Church. One way to
achieve this is for the laity to become deeply involved in the CLCC process, a
process that forms them systematically in Evangelization, Discipleship and
Service.
It is the laity who usually conduct the formation sessions. The pastor, however,
must take the initiative. He take time to become familiar with the process,
gives it his blessing, and promises to promote it by preaching and including it
in the Prayers of the Faithful each Sunday. He allows little progress reports to
be place in the parish bulletin. These are reminders that help keep the CLCC
alive in the parishioners' minds. They help people decide to join an existing
SEFC or form new ones in the parish.

\section{How and Where to Begin}

Ideally action begins when the pastor and parish council select a group of top
men and women called the Matrix 12 (after the 12 Apostles) who are qualified to
organize and lead the CLCC process. The Matrix 12 becomes the first Small
Ecclesial Faith Community in the parish. It models itself on Mary, Mother of God
and of the Church. As she conceived and delivered Jesus to the world in
Bethlehem, so the Matrix 12 delivers the CLCC process into the parish. Being the
first Small Ecclesial Faith Community in the parish, they begin their formation
in the faith, a spiritual journey of growing in grace by study, prayer and
bonding. Simultaneously they prepare for their mission, to reach out to lapsed,
unchurched Catholics with the aim of helping them return to the faith.
5
After some 9 to 12 months the Matrix 12 is formed well enough to establish more
SEFCs, a second, third, fourth, tenth, twentieth and so on. Hundreds of
Catholics will experience religious formation they've never had before. All the
while they are mindful of their mission to reach out to the lapsed. Filled with
the light of truth, they sense the power of Jesus' words in their lives, ``I am
the way, the truth and the life, no one comes to the Father except through me.''
(John 14.4)(Emphasis added) They begin Celebrating Life as Catholic Christians.

\section{The CLCC Process of Formation}

The CLCC formation process is an imitation, or as nearly as possible a
duplication in the parish of Jesus' mission in Galilee. He carried it out 5
Stages. So too the CLCC process is carried out in 5 parallel stages. They are
the following:

\subsection{Stage 1. Call of Apostles and Disciples}

Jesus first called disciples and later chose the 12 Apostles. Most notably they
were ordinary men, fishermen, tax collectors and others who were looking for the
Messiah. They were attracted to his personality. Some perhaps were looking for a
career change. He said simply, ``Come after me, and I will make you fishers of
men.'' (Mt. 4.19)

\subsection{Stage 2. Formation of the Apostles and Disciples}

Jesus' school of evangelization was not a classroom but the great outdoors: the
Sea of Galilee, a mountain, a vineyard, a field, a well-traveled road. He also
evangelized at a wedding feast, a banquet hall, a tax collector's house, the
home of Mary and Martha, the Temple and finally the Cross on Calvary and the
empty tomb. ``They were spellbound at his teaching because he spoke with
authority.'' (Lk.4.32)
6

\subsection{Stage 3. Stewardship of our Gifts}

The ``many and varied'' gifts we freely receive from God are in such abundance
and continuous flow that it is impossible for us to count them or even be aware
of them all. But as persons of faith we must ``count our blessings'' as best we
can. This means we must acknowledge God as the Giver of all gifts. It means
cultivating the habit of daily prayer to praise God for his gifts in justice,
thanking him humbly, rejoicing like the leper who was healed returned to Christ
``glorifying God in a loud voice.'' (Lk.17.15)

\subsection{Stage 4. Home Visitation and Evangelization for Community Building}

Jesus ``went around to the villages in the vicinity teaching. He summoned the
Twelve and began to send them out two by two and gave them authority over
unclean spirits.'' (Mk.6.6a-7) Jesus and the Twelve tried first to evangelize
the Chosen People. In Acts of the Apostles we read, ``the word of God be
spoken... first [to the Jews]... we now turn to the Gentiles.'' (Cf.13.46, also
47-48) Stage 4 in the CLCC process continues to form practicing Catholics more
deeply in faith and holiness, but now prepares them to turn to the thousands of
lapsed, non-church going, nominal Catholics residing within the parish
boundaries who are in need of salvation.

\subsection{Stage 5. Stewards of the Eucharist}

Most important in CLCC formation is Stage 5, the climax of the process. To be a
Catholic Christian one cannot get by with only a meager understanding of the
mystery of the Holy Eucharist, the ``greatest of all gifts.'' The Eucharist is
Jesus Christ himself, the Incarnate Word, The God Man, who is really and truly
present to man in his body and blood, soul and divinity. God the Father gave man
the supreme gift of himself, his only- begotten Son, Jesus, out of his infinite,
inestimable love of man. In turn Jesus gave himself to us. Before his ascension
into heaven, he chose to remain present to man out of love in a manner never
7
known before. Under the appearance of common bread and wine, eatable and potable
substances, he gave his Body and Blood, a reality known and believed only
through the gift of faith. By this unsurpassable gift, Catholics are given
access to union with Jesus Christ unlike any other union on earth. Thus, we must
regard ourselves as extraordinarily privileged ``servants of Christ and stewards
of the mysteries of God,'' as St. Paul says. (1 Cor.4.1) Note that first
encyclical of Pope Benedict XVI is titled God is Love.

\section{Formation is Weekly}

Formation sessions in the CLCC process are held once a week, every week
throughout the year. The success of this process is there are no interruptions
in the sequence. There are no lengthy intervals between get-togethers, for
example the months between Lent and autumn, no vacation breaks in the summer or
during the Christmas and Easter seasons or during the World Series. Continuity
keeps a steady flow of grace to all members. Their love of Christ and his Church
grow steadily, and bonding within the group is a sure sign of the presence and
influence of the Holy Spirit upon each individual. In view of experience with
the process, it is most strongly urged that adherence to the weekly session be
held ``sacred'' and not abandoned or yielded to personal feelings or easy
excuses. Realistically however, when a member of the SEFC must be absent for a
good reason, the weekly formation nonetheless goes on. Celebrating Life as a
Catholic Christian is perennial. It is not like a seasonal sport.

\section{Interior Spiritual Formation}

The weekly formation session has two parts. The first part is called Interior
Spiritual Formation, the second part is called Exterior Action. The first part
is a teaching, a presentation of key quotations from official Church documents
such as Vatican Council II, papal encyclicals, apostolic exhortations and
letters, the Catechism of the Catholic Church and related texts of Sacred
Scripture. The quotations are chosen in a way that their meaning and relevance
to the Stage under consideration is clear to all in the SEFC. The group is asked
to look carefully at the phrases and
8
important words in the paragraph, and center on one or two such phrases or words
invoking the Holy Spirit for light to get a hold on the main point. To be
avoided is to try to absorb the whole content thereby getting lost in a
multiplicity of thoughts and words that distract a person from the
teaching. Imperative also is that only one paragraph is read, studied, prayed
over and shared in the session. Reflection on one paragraph only at each session
is unique to this formation process. The insights the members obtain from the
paragraph and share with the group are always surprising and inspiring. Most
important to add here is that their process ought NOT be rushed through,
streamlined, shortened or gotten through with quickly. Rather each paragraph
must be read carefully, reflected on, savored, digested and shared in order to
draw from the knowledge and graces that only the Holy Spirit offers and only God
can count. Remember the majority of paragraphs are summary teachings of the
official Church. Therefore they have the power of truth derived from the Holy
Spirit. Cherish what you read, share what you have been given and receive what
you hear form the others as gifts of the Holy Spirit.
To help understand the one paragraph per week procedure and to gain maximum
spiritual formation from each session, at the end of each stage is a set of
Reflection and Discussion Issues (RDI) that correspond to the paragraph under
consideration. It is most helpful to refer to and use the RDIs. They are offered
to stimulate reflection. At the same time pray to the Holy Spirit for his gifts
of understanding and knowledge.

\section{Exterior Action, Part Two of CLCC Formation}

The second part of the weekly session is called Exterior Action. The paragraph
under Interior Spiritual Formation is reread and prayerfully reflected on again
in quiet time, but from the perspective of action: how can we act on what we
find in the paragraph? What worthwhile action can be chosen that translates the
mental process of reflection into a concrete missionary action that will help us
prepare for our outreach apostolate to fallen- away Catholics? Often ideas fly
freely. But without fail someone in the SEFC comes up with an action that is
acceptable to all. Not
9
infrequently the agreeable action is fine-tuned with someone offering a related
idea. The Exterior Action agreed upon becomes the action of the week for
all. Obviously the Holy Spirit is present and working. He is helping each person
in the SEFC to move forward on his or her journey of Celebrating Life as a
Catholic Christian.

\section{90 Minute Syndrome}

What makes the weekly CLCC formation session successful? The strict 90minute
time frame within which the weekly session is conducted! Packed into 90 minutes
are the two parts: 45minutes for Part I Interior Spiritual Formation, and 45
minutes for Part II Exterior Action. Each 45minute period follows a prescribed
order titled:

\section{Order for Conducting Formation Sessions of Small Ecclesial Faith Communities}

This Order is found at the beginning of the CLCC Manual. It is the solemn duty
of the Moderator, usually the leader of the SEFC, to follow faithfully this
Order. As members grow familiar with the procedure, other members are invited,
in fact urged, to take a turn at moderating a session. Moderating a session
gives each member a sense of belonging , and heightens his own participation
within the group. The experience gives everyone a feeling of responsibility for
the process and to the group.
The moderator must always stick to the Order and not deviate nor let others get
off track. He or she must be alert, wise, kind and strong. This means the
moderator may not allow people to wander off to other subjects not being
considered. Everyone should be encouraged to speak up and share. Should someone
get long- winded, the moderator with charity and prudence must cut off the
person in order to stick to the subject and stay within the time frame.
10
With respect to each person's circumstances it is most important that the
moderator starts the session on time and ends on time. A sure way to kill the
process is to start late or go overtime. After the session is over, if people
wish to stay to socialize that is their choice.
Where are the weekly meetings held? A room in the parish facility seems like a
good place because the SEFC then identifies with the parish. But it's also not a
good place. Other groups in the parish have a right to use the room too. The
better place is someone's private home. Someone in the SEFC is usually willing
to have the group of up to 12. This is how the early, apostolic Church startedin
peoples' homes.

\section{``By their fruits will you know them''}

To help the Matrix 12, the lay leadership group of the process, establish as
many Small Ecclesial Faith Communities as possible in the parish, it is useful
for them and those trying to decide whether to enter a SEFC to ponder the
parable of the ``good tree'' that bears ``good fruit.'' (Cf.Lk.6.43) The
application is easy to see. People who are thinking about entering the process,
that is, people desirous of growing in holiness and deeper knowledge of the
faith, can be likened to a ``good tree'' that bears ``good fruit.'' Already the
blossoms of faith begin to show. The buds of Christ- centered holiness break
forth. The tiny fruit, Celebrating Life as a Catholic Christian, is ripening to
become good fruit.
It is logical that a good tree bear good fruit. St. Luke writes, ``Every tree is
known by its fruit.'' (Cf.Lk.6.43-45) People who enter the CLCC formation
process are like the good tree that produces good fruit. The ``good fruit'' is
the many baptized Sunday Catholics, who have been minimally catechized, are
hungering for truth. By entering the process they feel they will ripen into
formed Catholic Christians. Their ``fruit'' is twofold: holiness and zeal to
spread the good news within, and outwardly
11
they want to build up the House of God by brining lapsed Catholics back into the
structure. Jesus said this, ``For every tree is known by its own fruit.'' (v.44)
St. Matthew narrates Jesus' fruit parable in this way, ``A good tree cannot bear
bad fruit, nor can a rotten tree bear good fruit. Every tree that does not bear
good fruit will be cut down and thrown into the fire. So by their fruits you
will know them.'' (Mt.7.18-20) (Emphasis added) Simply stated, the fruit of the
CLCC process is that people in parishes begin celebrating their lives as
Catholic Christians.
St. John in his Gospel records Jesus' parable of the vine, the vine grower,
branches and the fruit, an analogy that can be applied to the CLCC formation
process. It brings a powerful message to those in the process as well as to
those not yet in the process.
Jesus said, ``I am the vine, and my Father is the vine grower... Remain in me,
as I remain in you. Just as a branch cannot bear fruit on its own unless it
remains on the vine, so neither can you unless you remain in me. I am the vine,
you are the branches. Whoever remains in me and I in him will bear much fruit,
because without me you can do nothing.''
``It was not you who chose me, but I who chose you and appointed you to go and
bear fruit that will remain. So whatever you ask the Father in my name he may
give you.'' (Jn.15.1,4,5 and 16)
12

\section{Conclusion}

CLCC. This four-letter acronym has been used may times in this booklet. It
describes a formation process that every member of the Church will find
life-changing spiritually and intellectually rewarding if he chooses to enter
it.
The letters mean:
C Celebrating: God's grace together with the individual's strong faith make life
a joy, a jubilation
despite tribulation.
L Life: Because God created us to his image, and in baptism Christ recreated us
to his likeness, our lives are infinitely precious in his sight.
C Catholic: A unique identity has been given us. We are members of Christ's
Body. We belong to the
one, holy, catholic and apostolic Church that the Lord Jesus founded.
C Christian: As true followers of Christ we are committed totally to his way of
life. (Cf.Acts 11.26)

% --------------------------------------------------------------------

\end{document}                 % The input file ends with this command.
